The Shape Matching Element Method for Meshless Animation of NURBs Models

### 2021-01-19 paper/sections/methods.tex
ok remixed the methods so equations are where I think they should go, now need to make it make sense again

\section{Methods}
Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$. Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step. We note that we used a simplified version of D-NURBS in which the weights are held constant throughout the simulation and find this not to be limiting \myworries{(feel like we'll need better justification than saying just this?)}. We represent the degrees of freedom for the $i$-th NURBS surface with $m$ control points as $\mathbf{q}_i = \left[ \mathbf{p}_1^T, \dots, \mathbf{p}_m^T \right]^T \in \mathbb{R}^{3m}$. From this definition, we write the DOF for the entire model as $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ where $n$ is the number of NURBS surfaces for the model.

To evaluate the volumetric integrals in solid mechanics models, we require material points $\mathbf{X} \in \mathbb{R}^{3m}$ in the undeformed space that serve as integration points. Our shape matching element method enables us to evaluate the deformed positions of the material points strictly using positions defined on the boundary. Like in VEM, we fit a polynomial to each boundary primitive (NURBS in this case) describing its deformation. The fitting of these polynomials closely follows that of the Shape Matching \cite{10.1145/1073204.1073216} approach where we perform a least squares polynomial fit to the deformed boundary points. These boundary points are represented by sampled positions on each NURBS surface. Since each surface is associated with a single polynomial, we can construct a unique polynomial for some arbitrary point by blending these polynomials. Thus given some material point $\mathbf{X}_i$ we find it's corresponding deformed position $\mathbf{x}_i$ via a weighted combination of the polynomials. This permits us to build a general deformation map which we use to construct our equations of motion.

In the following sections, we will first review the simplified D-NURBS model used to form our generalized coordinates. We then present our shape matching algorithm for fitting polynomials to surfaces .... In section ... we next describe how to blend these surface polynomials to construct a general deformation map for material points. Section ... describes our strategy for selecting integration points in the volume through a raycasting approach. Finally, we formulate our equations of motions from a Lagrangian formulation.

\section{D-NURBS Model}
\myworries{TODO: Most of these matrices should be paranthesis matrices}
We provide a brief overview of the D-NURBS formulation presented in \cite{10.1145/176579.176580}. Our models in our method are composed of a collection of NURBS surfaces.

\subsection{D-NURBS Formulation}
A point on a NURBS surface is typically expressed in the following form
\begin{equation}
\label{eqn:nurbs_srf}
    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    {\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    \text{,}
\end{equation}
where where have a total of $nm$ control points $\mathbf{p}$ and weights $w$. Using the B-spline basis functions and the weighted control points, this formula gives a map from our parametric coordinates $u,v$ to their corresponding position on the surface in $\mathbb{R}^{3m}$. $B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.

@@

This formulation is a simplification of the full D-NURBS described in \cite{10.1145/176579.176580} the weights of the NURBS are fixed throughout the simulation. Permitting the dynamics to modify weights would likely improve the expressiveness of the kinematics, but we find fixing the weights affords sufficiently expressive results and improves performance. The performance improvement is a result of having fixed $N_{i,j}(u,v)$ entries over the course of the simulation. The full D-NURBS model requires rebuilding the jacobian and consequently the mass matrices at each time step.

\subsection{Generalized Coordinates for a NURBS model}
The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface. In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface. We emphasize that we only require an initial selection of $(u,v)$ coordinates across each surface and the dynamics of the simulation modifies the set of control points $\mathbf{q}$ to modify the world space maps $\mathbf{x}(u,v)=\mathbf{J}(u,v)\mathbf{q}$.

Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface. For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
\begin{equation}
\label{eqn:surface_jacobian}
    \mathbf{J}_i =
    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
    \text{.}
\end{equation}
Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$. Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
\begin{equation}
\mathbf{J} = \left[ \begin{array}{ccccc}
\mathbf{J}_1 &  &  &  &  \\
 & \ddots &  &  &  \\
 &  & \mathbf{J}_i & &  \\
 &  &  & \ddots &  \\
 &  &  &  & \mathbf{J}_n \\
\end{array} \right]
\text{.}
\end{equation}
With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.

\section{Shape Matching}
The key feature of our method is its unique deformation map $\phi(\mathbf{X}) = \mathbf{x}$ mapping some arbitrary undeformed material point $\mathbf{X}$ to its corresponding deformed position in space $\mathbf{x}$. We develop this similarly to the VEM approach in that we say the function we intend to reproduce may be some highly nonlinear one, but it may be approximated accurately by a polynomial. Thus we seek some $\phi$ function where  $\phi^* \approx \phi$ where $\phi^*$ is the true solution. In $1$-D this means we wish to find a function of the form $\phi(X) = c_1 + c_2X + c_3X^2 + \dots + c_nX^{n-1}$ where each $c_i$ is an unknown polynomial coefficient and the $X^i$ terms are elements of the polynomial basis.

\subsection{The Shape Matching Element}
In typical FEM in graphics, we make use of tetrahedral or triangular elements with a piecewise-linear function space. In such a case our functions may be evaluated as a weighted combination of the nodal values of an element: $f(\mathbf x)=\sum_{i=0}^n f_i \phi_i(\mathbf{x})$ where $n$ is the number of nodal values and $\phi_i$ is the i-th barycentric coordinates. Thus in this case we have the simple deformation map $\mathbf{x(X)} = \sum_{i=0}^n \mathbf{x}_i \phi_i(\mathbf{X})$. However, our method contrasts this typical approach in that our elements are more general and take the form of any arbitrary boundary primitive, such as NURBS, and also that the function space in non-linear.

In our algorithm, a single shape matching element is represented by a NURBS surface, a center of mass, and polynomial describing its current deformation. We construct these polynomials via a shape matching strategy. With the addition of the center of mass, we describe the shapes deformation about this point and thus the monomial basis in the case of a linear polynomial is $\left[ \mathbf 1^T (\mathbf{X - \bar{X}})^T \right]$ where $\mathbf 1^T$ is a constant $1$ column vector and $\mathbf{\bar{X}}$ is the center of mass for the element. To simplify the following formulas, we collect the non-constant terms of the monomial basis into a single vector $\mathbf{m}(\mathbf{X}-\mathbf{\bar{X}})$. For example if $\mathbf{X} = (X,Y,Z)$, $\mathbf{\bar{X}} = (\bar X, \bar Y, \bar Z)$, and $\mathbf{q}=\mathbf{X}-\mathbf{\bar{X}}$ with a quadratic polynomial we would have
\begin{equation}
m(\mathbf{q)} = (
					q_1, q_2, q_3,
                   	q_1^2, q_2^2, q_3^3,
                   	q_1 q_2, q_2 q_3
                ) \in \mathbb{R}^k
                \text{,}
\end{equation}
where $k$ is the size of the monomial basis. Next we arrange this basis into a matrix of the form
\begin{equation}
\mathbf{M}(\mathbf{q}) = \left[ \begin{array}{ccc}
\mathbf{m}(\mathbf{q}) &  &  \\
 & \mathbf{m}(\mathbf{q})  &  \\
 &  & \mathbf{m}(\mathbf{q})  \\
\end{array} \right] \in \mathbb{R}^{3x3k}
\text{.}
\end{equation}
With this definition of the monomial basis, we may reconstruct the deformed position $\mathbf{x}$ for some material point $\mathbf{X}$ with
\begin{equation}
\mathbf{x}(\mathbf{X}) = \left[\mathbf{M}(\mathbf{X-\bar{X}}) \quad \mathbf{I} \right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right]
\text{,}
\end{equation}
where $\mathbf{c} \in \mathbb{R}^{3k}$ is the set of polynomial coefficients for the non-constant terms, $\mathbf{t} \in \mathbb{R}^3$ is the constant term coefficients, and $\mathbf{I} \in \mathbb{R}^{3x3}$ is an identity matrix. The reason for separating out the constant term becomes more evident when we examine case where we have multiple elements.

@@

\subsection{Shape Matching with Multiple Elements}
In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface. Therefore for the entire model each shape has a boundary, a polynomial describing its deformation, as well as an associated center of mass. These centers of mass may not be distinct, meaning that a single center of mass may be shared among multiple surfaces. This sharing of center of masses is typical and in some cases required (\myworries{reference center of mass section that describes how for planar surfaces we need to share COM to get a COM out of plane}). This leads to a new definition of the center of mass where it is the mean of all the boundary points associated with it.

@@

\subsection{Hierarchical fitting}
A problem may emerge in the shape matching process where the system may be underdetermined. To account for this we solve for the coefficients in a hierarchical manner where we initially solve for the linear terms and at each step solve the coefficients for the next higher order of deformation. The produces a set of smaller systems which are guaranteed to be determined \myworries{probably need to justify this more}. The intuition behind this can be described by the \textit{hierarchical ordering principle}, which argues that the behavior of a system is typically dominated by lower order effects \myworries{cite something here.}. To describe our fitting strategy, we begin with the following shape matching system for a single element
\begin{equation}
\label{eqn:fitting_one_copy}
\left[
\mathbf{A} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{.}
\end{equation}
We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$. Similarly, we also have $\mathbf{A}_{\text{lin}}$ and $\mathbf{A}_{\text{quad}}$ which are the blocks of $\mathbf{A}$ corresponding to the columns for the linear and quadratic monomial terms, respectively. First we construct a system for the linear terms:
\begin{equation}
\label{eqn:fitting_linear}
\left[
\mathbf{A}_{\text{lin}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{lin}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{,}
\end{equation}
which leads to the following expression for the linear coefficients
\begin{equation}
\label{eqn:c_linear}
\begin{split}
\mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
           & = \mathbf{L}_{\text{lin}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
\begin{equation}
\label{eqn:fitting_quadratic}
\left[
\mathbf{A}_{\text{quad}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{quad}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
\text{.}
\end{equation}
Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
\begin{equation}
\label{eqn:c_quad}
\begin{split}
\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
           & = \mathbf{L}_{\text{quad}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
The final form for the full $\mathbf{L}$ matrix is thus given by
\begin{equation}
\mathbf{L} =
\begin{bmatrix}
\mathbf{L}_{\text{lin}} \\
\mathbf{L}_{\text{quad}} \\
\mathbf{T}
\end{bmatrix}
\end{equation}

@@

\section{The Deformation Map}
At this point we have shown how to fit a polynomial describing the deformation of each surface. With these polynomial coefficients coupled with their associated centers of mass, we can construct an estimate of some arbitrary material point $\mathbf{X}$ using the deformation described by the surface. This lends itself to a general definition for the deformed position of a point as a weighted sum of the polynomials, which yields a unique polynomial for the point. Thus we arrive at a definition for the deformed position, $\mathbf{x}$, of a material point $\mathbf{X}$ in the undeformed space:
\begin{equation}
\label{eqn:deformation_map}
	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
	\text{,}
\end{equation}
where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively. The terms $\mathbf{\bar{X}}_j$ and $\mathbf{t}_j$ the undeformed and deformed centers of mass, respectively, that are associated with the $i$-th surface.

\subsection{Computing the Blending Weights}
All that remains for the deformation map is to compute blending weights. Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest. This idea lends itself to a simple distance-based weighting method. The first step in this process is to compute \textit{distance weights}. For a given point $\mathbf{X}$ and a set of $n$ we wish to compute a set of distance weights to each surface $\mathbf{\theta}(\mathbf{X}) = \left(\theta_1, \dots, \theta_i, \dots, \theta_n \right)$.

@@

\section{Selecting Centers of Mass}
\myworries{TODO. waiting to implement this before I start writing how we do it}
\section{Raycasting Quadrature}
The last step before we may discuss the dynamic simulation is to describe our strategy for generating material points $\mathbf{X}$ in the undeformed model. Our volume integration strategy is based on \cite{KHOSRAVIFARD201030} which we briefly describe in the following section
\subsection{Review of the Meshless Method for Domain Integral Evaluation in \cite{KHOSRAVIFARD201030}}
This paper addresses the integration of volume integrals of the form $\mathbf{I} = \int_\Omega f(\mathbf{x}) d\Omega$ where $\mathbf{x} = (x_1,x_2,x_3)$. The authors show that through application of the Green-Gauss theorem, this integral may be rewritten as 2D integral over a $yz$ plane where at each position along the plane we evaluate a line integral along the $x$-axis. To evaluate these line integrals, a raycasting procedure is performed to find a set of intersection along the line. These intersection define integration domains in which we sample 3D integration points. As a result of integrating over the 2D plane and raycasting at each 2D point, we produce an integration rule over the entire domain that can converge to the true integral value.

@@

\section{Dynamic Simulation}
With our definition of the deformation map as well as a strategy for integrating the volume, we now have the necessary ingredient to perform dynamic simulation. We follow the Variational Mechanics approach that defines a pair of kinetic and potential energies, and we apply the Euler-Lagrange equation to develop our equations of motion.

\subsection{Generalized Inertia}

@@

\subsection{Generalized Forces}
We write our potential energy as 
    V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
        \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
               = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$. Full details on how to compute $\mathbf{F}$ and its derivative $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ may be found in the appendix. (\myworries{I just wrote this because I'm too tired to right the d etails right now :p}. Now that we have an expression for the deformation gradient, we discretize our potential energy as follows
    V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.

Like in the case of the kinetic energy, we also introduce an error force potential, which takes the simple form
\begin{equation}
V_E = \mathbf{q}^T \mathbf{M}_E \mathbf{q}
\end{equation}

\subsection{Time Integration}
We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
\begin{equation}
    L=(T+T_E)-(V + V_E)
\end{equation}
Then from the Euler-Lagrange equation we have:
\begin{equation}
    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
\end{equation}
\myworries{todo: not totally sure what to write here since we're experimenting with both linearly implicit backward euler as well as Newton's method soon}

>>>>

\section{Overview}
The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
(density, constitutive model, model parameters etc). The output is an elastodynamic simulation. 
SEM consists of preprocessing and runtime simulation phases~\refalg{sem} and at no point do we need to generate a volumetric mesh of the input. 
In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
Finally we construct local shape matching operators and construct the mass matrix for our problem.

At runtime we use standard time integration schemes to time step the system (we show linearly implicit, fully implicit, runge-kutta four). 
These require us to evaluate the energy, gradient and Hessian (implicit only) which we do at the previously computed quadrature points. 
For collisions we use penalty forces and apply them as external forces during integration.

\begin{algorithm}[h]
    \label{alg:sem}
    \SetAlgoLined
    \KwResult{Write here the result }
     initialization\;
     \While{While condition}{
      instructions\;
      \eIf{condition}{
       instructions1\;
       instructions2\;
       }{
       instructions3\;
      }
     }
     \caption{SEM Algorithm}
    \end{algorithm}

\section{Methods}

\dave{explain NURBS but not D-NURBS because that's Terzopolous's specific algorithm just explain the NURBS parameterization}

A point on a NURBS surface is typically expressed in the following form
\begin{equation}
\label{eqn:nurbs_srf}
    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    {\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    \text{,}
\end{equation}
where where have a total of $nm$ control points $\mathbf{p}$ and weights $w$. Using the B-spline basis functions and the weighted control points, this formula gives a map from our parametric coordinates $u,v$ to their corresponding position on the surface in $\mathbb{R}^{3m}$. $B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.

@@

With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.

\begin{figure}
    \includegraphics[width=\columnwidth]{example-image-a}
    \caption{Figure showing input model NURBS parts and shape matching to each part}
    \label{fig:shapematching}
\end{figure}

\subsection{Shape Matching}
\dave{Shape Matching to a single NURBS part}
In typical FEM in graphics, we make use of tetrahedral or triangular elements with a piecewise-linear function space. In such a case our functions may be evaluated as a weighted combination of the nodal values of an element: $f(\mathbf x)=\sum_{i=0}^n f_i \phi_i(\mathbf{x})$ where $n$ is the number of nodal values and $\phi_i$ is the i-th barycentric coordinates. Thus in this case we have the simple deformation map $\mathbf{x(X)} = \sum_{i=0}^n \mathbf{x}_i \phi_i(\mathbf{X})$. However, our method contrasts this typical approach in that our elements are more general and take the form of any arbitrary boundary primitive, such as NURBS, and also that the function space in non-linear.

In our algorithm, a single shape matching element is represented by a NURBS surface, a center of mass, and polynomial describing its current deformation. We construct these polynomials via a shape matching strategy. With the addition of the center of mass, we describe the shapes deformation about this point and thus the monomial basis in the case of a linear polynomial is $\left[ \mathbf 1^T (\mathbf{X - \bar{X}})^T \right]$ where $\mathbf 1^T$ is a constant $1$ column vector and $\mathbf{\bar{X}}$ is the center of mass for the element. To simplify the following formulas, we collect the non-constant terms of the monomial basis into a single vector $\mathbf{m}(\mathbf{X}-\mathbf{\bar{X}})$. For example if $\mathbf{X} = (X,Y,Z)$, $\mathbf{\bar{X}} = (\bar X, \bar Y, \bar Z)$, and $\mathbf{q}=\mathbf{X}-\mathbf{\bar{X}}$ with a quadratic polynomial we would have

\begin{figure}
    \includegraphics[width=\columnwidth]{example-image-a}
    \caption{Figure showing shape matching to NURBS}
    \label{fig:shapematching}
\end{figure}

\begin{equation}
m(\mathbf{q)} = (
					q_1, q_2, q_3,
                   	q_1^2, q_2^2, q_3^3,
                   	q_1 q_2, q_2 q_3
                ) \in \mathbb{R}^k
                \text{,}
\end{equation}
where $k$ is the size of the monomial basis. Next we arrange this basis into a matrix of the form
\begin{equation}
\mathbf{M}(\mathbf{q}) = \left[ \begin{array}{ccc}
\mathbf{m}(\mathbf{q}) &  &  \\
 & \mathbf{m}(\mathbf{q})  &  \\
 &  & \mathbf{m}(\mathbf{q})  \\
\end{array} \right] \in \mathbb{R}^{3x3k}
\text{.}
\end{equation}
With this definition of the monomial basis, we may reconstruct the deformed position $\mathbf{x}$ for some material point $\mathbf{X}$ with
\begin{equation}
\mathbf{x}(\mathbf{X}) = \left[\mathbf{M}(\mathbf{X-\bar{X}}) \quad \mathbf{I} \right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right]
\text{,}
\end{equation}
where $\mathbf{c} \in \mathbb{R}^{3k}$ is the set of polynomial coefficients for the non-constant terms, $\mathbf{t} \in \mathbb{R}^3$ is the constant term coefficients, and $\mathbf{I} \in \mathbb{R}^{3x3}$ is an identity matrix. The reason for separating out the constant term becomes more evident when we examine case where we have multiple elements.

\dave{here explain sampling on NURBS surface to get material points}
The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface. In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface. We emphasize that we only require an initial selection of $(u,v)$ coordinates across each surface and the dynamics of the simulation modifies the set of control points $\mathbf{q}$ to modify the world space maps $\mathbf{x}(u,v)=\mathbf{J}(u,v)\mathbf{q}$.

Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface. For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
\begin{equation}
\label{eqn:surface_jacobian}
    \mathbf{J}_i =
    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
    \text{.}
\end{equation}
Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$. Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
\begin{equation}
\mathbf{J} = \left[ \begin{array}{ccccc}
\mathbf{J}_1 &  &  &  &  \\
 & \ddots &  &  &  \\
 &  & \mathbf{J}_i & &  \\
 &  &  & \ddots &  \\
 &  &  &  & \mathbf{J}_n \\
\end{array} \right]
\text{.}
\end{equation}

@@

\subsection{Shape Matching Element Method}
\dave{Multiple Parts}
\dave{each polynomial/NURBS tuple is a shape matching element}
\dave{Shape Matching Element assembles global kinematics using blendeded formultion}
 In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface. Therefore for the entire model each shape has a boundary, a polynomial describing its deformation, as well as an associated center of mass. These centers of mass may not be distinct, meaning that a single center of mass may be shared among multiple surfaces. This sharing of center of masses is typical and in some cases required (\myworries{reference center of mass section that describes how for planar surfaces we need to share COM to get a COM out of plane}). This leads to a new definition of the center of mass where it is the mean of all the boundary points associated with it.

@@

\paragraph*{Hierarchical fitting}
A problem may emerge in the shape matching process where the system may be underdetermined. To account for this we solve for the coefficients in a hierarchical manner where we initially solve for the linear terms and at each step solve the coefficients for the next higher order of deformation. The produces a set of smaller systems which are guaranteed to be determined \myworries{probably need to justify this more}. The intuition behind this can be described by the \textit{hierarchical ordering principle}, which argues that the behavior of a system is typically dominated by lower order effects \myworries{cite something here.}. To describe our fitting strategy, we begin with the following shape matching system for a single element
\begin{equation}
\label{eqn:fitting_one_copy}
\left[
\mathbf{A} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{.}
\end{equation}
We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$. Similarly, we also have $\mathbf{A}_{\text{lin}}$ and $\mathbf{A}_{\text{quad}}$ which are the blocks of $\mathbf{A}$ corresponding to the columns for the linear and quadratic monomial terms, respectively. First we construct a system for the linear terms:
\begin{equation}
\label{eqn:fitting_linear}
\left[
\mathbf{A}_{\text{lin}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{lin}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{,}
\end{equation}
which leads to the following expression for the linear coefficients
\begin{equation}
\label{eqn:c_linear}
\begin{split}
\mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
           & = \mathbf{L}_{\text{lin}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
\begin{equation}
\label{eqn:fitting_quadratic}
\left[
\mathbf{A}_{\text{quad}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{quad}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
\text{.}
\end{equation}
Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
\begin{equation}
\label{eqn:c_quad}
\begin{split}
\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
           & = \mathbf{L}_{\text{quad}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
The final form for the full $\mathbf{L}$ matrix is thus given by
\begin{equation}
\mathbf{L} =
\begin{bmatrix}
\mathbf{L}_{\text{lin}} \\
\mathbf{L}_{\text{quad}} \\
\mathbf{T}
\end{bmatrix}
\end{equation}

@@

\subsection{Blending Weights}
\dave{Desireable Properties}
\dave{Construction via post normalization}
\dave{only need blending weights at quadrature points and surface samples so given sparse set of quadrature points compute blending weights using raycasting}
\dave{things to explain (1) raycasting works for overlapping geometric, seperated geometry (2) weights glue objects together}
\dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
All that remains for the deformation map is to compute blending weights. Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest. This idea lends itself to a simple distance-based weighting method. The first step in this process is to compute \textit{distance weights}. For a given point $\mathbf{X}$ and a set of $n$ we wish to compute a set of distance weights to each surface $\mathbf{\theta}(\mathbf{X}) = \left(\theta_1, \dots, \theta_i, \dots, \theta_n \right)$.

@@

\subsection{Multiple Part Projection}
\dave{explain why you need to share constant terms (fit polynomial around a point in space), explain clustering and how to pick, forward reference to quadrature point}
\subsection{Quadrature Points}
\dave{quadrature via raycasting}
The last step before we may discuss the dynamic simulation is to describe our strategy for generating material points $\mathbf{X}$ in the undeformed model. Our volume integration strategy is based on \cite{KHOSRAVIFARD201030} which we briefly describe in the following section

This paper addresses the integration of volume integrals of the form $\mathbf{I} = \int_\Omega f(\mathbf{x}) d\Omega$ where $\mathbf{x} = (x_1,x_2,x_3)$. The authors show that through application of the Green-Gauss theorem, this integral may be rewritten as 2D integral over a $yz$ plane where at each position along the plane we evaluate a line integral along the $x$-axis. To evaluate these line integrals, a raycasting procedure is performed to find a set of intersection along the line. These intersection define integration domains in which we sample 3D integration points. As a result of integrating over the 2D plane and raycasting at each 2D point, we produce an integration rule over the entire domain that can converge to the true integral value.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{example-image-a}
    \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}
    \label{fig:raycasting}
\end{figure}

\subsection{Mass Matrix}

@@

% \subsection{Generalized Forces}
% We write our potential energy as 
% \begin{equation}
%     V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
% \text{.}
% \end{equation}
% To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
% \begin{equation}
%         \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
%                = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
% \text{.}
% \end{equation}
% In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$. Full details on how to compute $\mathbf{F}$ and its derivative $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ may be found in the appendix. (\myworries{I just wrote this because I'm too tired to right the d etails right now :p}. Now that we have an expression for the deformation gradient, we discretize our potential energy as follows
% \begin{equation}
%     V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
% \text{,}
% \end{equation}
% where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.
